---
title:  "CA-WRF-regional-runs"
author: "Xiulin Gao"
email:  "xiulingao@lbl.gov"
date:   "10-Nov-2022"
---
## Introduction
This script is used to process regional runs in California grasslands. 
We want to benchmark model simulations of annual burned fraction, GPP, and LAI in annual grasslands. 
This scripts uses a set of scripts developed by Marcos Longo at https://github.com/mpaiao/FATES_Utils;
part of this script is also adapted from Marcos Longo's post-processing script, changes are
made to do post-processing specifically for 1PFT regional simulations. 



```{r, label = 'reset-R',message=FALSE, results='hide'}
# Unload all packages except for the R default ones
plist = names(sessionInfo()$otherPkgs)
if (length(plist) > 0){
   dummy = sapply(X=paste0("package:",plist),FUN=detach,character.only=TRUE,unload=TRUE)
}#end if (length(plist) > 0)


# Remove all variables
rm(list=ls())

# Reset warnings
options(warn=0)

# Close all plots
invisible(graphics.off())

# Clean up
invisible(gc())
```






```{r, label="path"}
home_path   = path.expand("~")
local_path  = file.path(home_path,"Downloads/regional-runs")
main_path   = file.path("~/Google Drive/My Drive/all")
region_path = file.path(main_path,"regional-runs")
util_path   = file.path(home_path,"Util/RUtils")


cases       = c( "regional-low-genl-N05T08_CLM_FATES.clm2.nc"
               , "regional-mean-brdi-N09T04_CLM_FATES.clm2.nc"
               , "regional-median-genl-N12T18_CLM_FATES.clm2.nc"
               , "regional-quant-genl-N08T27_CLM_FATES.clm2.nc"
              )

case_files   = file.path(local_path,cases)
ncase        = length(cases)
case_names   = c(1L,2L,3L,4L)

case_desc    = c( "Low-GENL-G1-N05T08"
               , "Mean-BRDI-G3-N09T04"
               , "Median-GENL-G1-N1218"
               , "Quant-GENL-G1-N08T27"
               )

names(case_desc) = c("1","2","3","4")

#Observation path
obs_main    = file.path("~/Google Drive/My Drive/CA-grassland-simulationDoc/benchmark")
site_main   = file.path(obs_main,"Vaira-Ranch-AGB-LAI")
pwood_main  = file.path(obs_main,"pepperwood-biomass-data")
seablm_main = file.path(obs_main, "Seabloom-CA-grassland-productivity")
frap_main   = file.path(obs_main,"FRAP")
sif_main    = file.path(obs_main,"SIF-GPP-global0.05degree-monthly")
lai_main    = file.path(obs_main,"LAI-monthly-pt05degree")
site_base   = list.files(path=site_main,pattern="_eddy-summ\\.nc$"  )
site_file   = file.path(site_main,site_base)
site_agb    = file.path(site_main,list.files(path=site_main,pattern="biomass"))
site_lai    = file.path(site_main,list.files(path=site_main,pattern="LAI\\.csv$"))
agb_base    = list.files(path=pwood_main,pattern=".csv")
#lai_base    = list.files(path=lai_main,pattern="LAI\\.csv$")
agbS_path   = file.path(seablm_main,"nutnet-npp-div-data-output.csv")
#flux_path   = file.path(sif_main,flux_base)
agbP_path   = file.path(pwood_main,agb_base)
#lai_path    = file.path(site_main,lai_base)
frap_grass  = file.path(frap_main,"Bfrac-WRF-grassonly.csv")
wrf_path    = file.path("~/Google Drive/My Drive/9km-WRF-1980-2020/1981-01.nc")

```






```{r,label="make-dir"}

#plot path
plot_main  = file.path(region_path,"Figures")
# Output path for time series
tsage_path    = file.path(plot_main,"tseries_age"         )
tsdbh_path    = file.path(plot_main,"tseries_dbh"         )
tspft_path    = file.path(plot_main,"tseries_pft"         )
tsmort_path   = file.path(plot_main,"tseries_mort"        )
tsnppo_path   = file.path(plot_main,"tseries_npp_organ"   )
tsauto_path   = file.path(plot_main,"tseries_auto_resp"   )
tsdap_path    = file.path(plot_main,"ts_heat_dbh+pft"     )
tsdphen_path  = file.path(plot_main,"ts_drought_phenology")
tssoil_path   = file.path(plot_main,"ts_heat_soil"        )
tstheme_path  = file.path(plot_main,"tseries_theme"       )
secy_path     = file.path(plot_main,"seasonal-cycle"      )
sens_path     = file.path(plot_main,"variance-decomp"     )

# Create paths for time series
dummy = dir.create(tsage_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsdbh_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tspft_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsmort_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsnppo_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsauto_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsdap_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsdphen_path, recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tssoil_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tstheme_path, recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(secy_path   , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(sens_path   , recursive = TRUE, showWarnings = FALSE)

```





```{r,label="user-pft-fuel"}
user_pftinfo  = TRUE
user_fuelinfo = TRUE

# List of PFTs, in case user_pftinfo is TRUE
n            = 0
pftinfo      = list()
n            = n + 1
pftinfo[[n]] = list( id     = 1
                   , key    = "pft1"
                   , short  = "Pooideae"
                   , desc   = "C3 drought decid grass"
                   , colour = "#00BFC4"
                   , parse  ="D*e*c*i*d[Non-h*y*d*r*o]"
                   )#end of PFT list


# List of fuels, in case user_fuleinfo is TRUE
n               = 0
fuelinfo        = list()
n               = n + 1
fuelinfo[[n]]   = list( id      = 1
                      , key     = "fuel1"
                      , short   = "Twig"
                      , desc    = "Fuel from twig" 
                      , colour  = "#E69F00"
                      , parse   = "F*u*e*l[T*w*i*g]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 2
                      , key     = "fuel2"
                      , short   = "SBranch"
                      , desc    = "Fuel from small branch" 
                      , colour  = "#56B4E9"
                      , parse   = "F*u*e*l[S*b*r*a*n*c*h]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 3
                      , key     = "fuel3"
                      , short   = "LBranch"
                      , desc    = "Fuel from large branch" 
                      , colour  = "#0072B2"
                      , parse   = "F*u*e*l[L*b*r*a*n*c*h]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 4
                      , key     = "fuel4"
                      , short   = "Trunk"
                      , desc    = "Fuel from trunk" 
                      , colour  = "#D55E00"
                      , parse   = "F*u*e*l[T*r*u*n*k]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 5
                      , key     = "fuel5"
                      , short   = "DeadLeaf"
                      , desc    = "Fuel from dead leaf" 
                      , colour  = "#999999"
                      , parse   = "F*u*e*l[D*l*e*a*f]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 6
                      , key     = "fuel6"
                      , short   = "LiveGrass"
                      , desc    = "Fuel from live grass" 
                      , colour  = "#009E73"
                      , parse   = "F*u*e*l[L*g*r*a*s*s]") #end of fuel list



```






```{r,label="plot-setting"}

gg_device  = c("png")     # Output devices to use (Check ggsave for acceptable formats)
gg_depth   = 600          # Plot resolution (dpi)
gg_ptsz    = 18           # Font size
gg_ptszl   = 26
gg_width   = 17.5         # Plot width (units below)
gg_widthn  = 14.5
gg_height  = 8.5          # Plot height (units below)
gg_units   = "in"         # Units for plot size
gg_screen  = TRUE         # Show plots on screen as well?
gg_tfmt    = "%y"         # Format for time strings in the time series %Y: 4 DIGITS YEAR %y 2 digits year
gg_ncolours    = 129      # Number of node colours for heat maps.
gg_fleg        = 1./6.    # Fraction of plotting area dedicated for legend
gg_dstat_thick = 0.1      # Thickess for the drought-deciduous status band in the stress overview plot (Value

#ens_colour  =  c("#E69F00", "#56B4E9", "#CC79A7",
                #"#F0E442", "#0000FF", "#00B286")

ens_colour = c("#A0AE6A", "#836B43", "#D68D18", "#437683", "#18B0D6")

top12 <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
           "#44AA99", "#999933", "#882255", "#6699CC", "#888888", "#661100")

site_colour  = "#000000"

ndevice = length(gg_device)


# Define lower and upper bound for interannual variability ribbon around the mean.
sdev_ribbon  = 1.                  # Standard-deviation equivalent for ribbon 
qlwr_ribbon  = pnorm(-sdev_ribbon) # Lower quantile for ribbon
qupr_ribbon  = pnorm(+sdev_ribbon) # Lower quantile for ribbon
alpha_ribbon = 0.2

```




```{r, label="load-everything"}
source(file.path(util_path,"load.everything.r"),chdir=TRUE)
library(stars)

```







```{r,label="time-location-setting"}

nc_conn      = nc_open(case_files[1])

# longitude and latitude
wrf         = nc_open(wrf_path)
lon         = ncvar_get(wrf, "LONGXY")
lat         = ncvar_get(wrf, "LATIXY")
dummy       = nc_close(wrf)
lon_vec     = as.vector(lon)
lat_vec     = as.vector(lat)
nlon        = length(lon_vec)
nlat        = length(lat_vec)

# time
time_val     = ncvar_get(nc_conn,"time")
time_use     = time_val[229:480] #only retrieve values for years during 2000-2020
nstep        = length(time_use)
simu_time    = c(2000:2020)
model_yra    = 1981
n_yr         = round((tail(time_use,1)-head(time_use,1))/365,digits=0)
yr_start     = model_yra + floor(time_use[1]/365)
yr_end       = yr_start + n_yr-1
nc_year      = as.numeric(yr_start:yr_end)
nc_month     = as.numeric(1:12)
yr_ens       = rep(nc_year,each=nlon*12)
month_yr     = rep(nc_month,each=nlon)
month_ens    = rep(month_yr,times=n_yr)
ens_tstamp   = make_datetime(year=yr_ens,month=month_ens,day=1L)
ntstamp_ens = length(ens_tstamp)
dummy        = nc_close(nc_conn)

```





```{r,label="dimension-setting"}
if ("nc_conn" %in% ls()){dummy = nc_close(nc_conn); rm(nc_conn)}

nc_dlist <- list()
nc_vlist <- list()

nc_conn  = nc_open(filename=case_files[1])
nc_nvars = nc_conn$nvars
nc_ndims = nc_conn$ndims
nc_dlist = rep(NA_character_,times=nc_ndims)
nc_vlist = rep(NA_character_,times=nc_nvars)
for (d in sequence(nc_ndims)) nc_dlist[d] = nc_conn$dim[[d]]$name
for (v in sequence(nc_nvars)) nc_vlist[v] = nc_conn$var[[v]]$name

#---~---
# Gather dimension information, then initialize matrices
#---~---

# List of age classes
idxage   = match("fates_levage",nc_dlist)
if (is.finite(idxage)){
   ages     = nc_conn$dim[[idxage]]$vals
   nages    = nc_conn$dim[[idxage]]$len

   ageinfo  = tibble( id      = sequence(nages)
                    , age_lwr = ages
                    , age_upr = c(ages[-1],Inf)
                    , key     = sprintf("age_%3.3i",ages)
                    , desc    = c( paste0("paste(paste(",age_lwr[-nages],"<= A*g*e)<",age_upr[-nages],"*y*r)")
                                 , paste0("paste( A*g*e >=",age_upr[nages],"*y*r)")
                                 )#end c
                    , labs    = c( paste0("paste(",age_lwr[-nages],"-",age_upr[-nages],")")
                                 , paste0("paste(",age_lwr[ nages],"-infinity)")
                                 )#end c
                    , colour  = viridis(nages,option="D",direction=-1)
                    )#end tibble
}else{
   ageinfo  = tibble( id      = integer(0L)
                    , age_lwr = numeric(0L)
                    , age_upr = numeric(0L)
                    , key     = character(0L)
                    , desc    = character(0L)
                    , labs    = character(0L)
                    , colour  = character(0L)
                    )#end tibble
}#end if (is.na(idxage))

# Set number of age classes
nages = nrow(ageinfo)


# List of size classes
idxdbh   = match("fates_levscls",nc_dlist)
if (is.finite(idxdbh)){
   dbhs     = nc_conn$dim[[idxdbh]]$vals
   ndbhs    = nc_conn$dim[[idxdbh]]$len
   dbhinfo  = tibble( id      = sequence(ndbhs)
                    , dbh_lwr = dbhs
                    , dbh_upr = c(dbh_lwr[-1],dbh_lwr[ndbhs]+2*max(diff(dbh_lwr)))
                    , dbh     = 0.5 * (dbh_lwr + dbh_upr)
                    , key     = sprintf("dbh_%3.3i",dbh_lwr)
                    , desc    = c( paste0("paste(paste(",dbh_lwr[-ndbhs],"<=D*B*H)<",dbh_upr[-ndbhs],"*c*m)")
                                 , paste0("paste( D*B*H >=",dbh_lwr[ndbhs],"*c*m)")
                                 )#end c
                    , labs    = c( paste0("paste(",dbh_lwr[-ndbhs],"-",dbh_upr[-ndbhs],")")
                                 , paste0("paste(",dbh_lwr[ ndbhs],"-infinity)")
                                 )#end dbhlabs
                    , colour   = magma(ndbhs,direction=1)
                    )#end tibble
}else{
   dbhinfo  = tibble( id      = integer(0L)
                    , dbh_lwr = numeric(0L)
                    , dbh_upr = numeric(0L)
                    , key     = character(0L)
                    , desc    = character(0L)
                    , labs    = character(0L)
                    , colour  = character(0L)
                    )#end tibble
}#end if (is.finite(idxdbh))

# Set number of size classes
ndbhs = nrow(dbhinfo)


# List of PFT classes (only if not using user-defined classes).
idxpft   = match("fates_levpft",nc_dlist)
if (! is.finite(idxpft)){
   # PFT index not found. Skip PFTs altogether.
   pftinfo = tibble( id               = numeric(0L)
                   , key              = character(0L)
                   , short            = character(0L)
                   , desc             = character(0L)
                   , colour           = character(0L)
                   , stringsAsFactors = FALSE
                   )#end data.table
}else if (! user_pftinfo){
   # Select all PFTs available
   pftids  = nc_conn$dim[[idxpft]]$vals
   npftids = nc_conn$dim[[idxpft]]$len

   # Build tibble with all the PFTs.
   pftinfo = tibble( id               = pftids
                   , key              = sprintf("pft%2.2i" ,pftids)
                   , short            = sprintf("PFT%2.2i" ,pftids)
                   , desc             = sprintf("PFT %2.2i",pftids)
                   , colour           = brewer.pal(n=npftids,name="PuBuGn")
                   , stringsAsFactors = FALSE
                   )#end tibble
}else if (! is_tibble(pftinfo)){
   # Convert user-defined pftinfo to a "tibble" object
   pftinfo  = do.call(what=rbind,args=lapply(X=pftinfo,FUN=as_tibble,stringsAsFactors=FALSE))
}#end if (! is.finite(idxpft))

# Set number of PFTs (active PFTs only) 
npfts = nrow(pftinfo)

# List of fuel classes

idxfuel   = match("fates_levfuel",nc_dlist)
if (! is.finite(idxfuel)){
   # fuel index not found. Skip Fuels altogether.
   fuelinfo = tibble( id               = numeric(0L)
                    , key              = character(0L)
                    , short            = character(0L)
                    , desc             = character(0L)
                    , colour           = character(0L)
                    , stringsAsFactors = FALSE
                    )#end data.table
}else if (! user_fuelinfo){
   # Select all fuels available
   fuelids  = nc_conn$dim[[idxfuel]]$vals
   nfuelids = nc_conn$dim[[idxfuel]]$len

   # Build tibble with all the fuels.
   fuelinfo = tibble( id               = fuelids
                    , key              = sprintf("Fuel %2.2i" ,fuelids)
                    , short            = sprintf("Fuel %2.2i" ,fuelids)
                    , desc             = sprintf("Fuel %2.2i" ,fuelids)
                    , colour           = brewer.pal(n=nfuelids,name="PuBuGn")
                    , stringsAsFactors = FALSE
                   )#end tibble
}else if (! is_tibble(fuelinfo)){
   # Convert user-defined fuelinfo to a "tibble" object
   fuelinfo  = do.call(what=rbind,args=lapply(X=fuelinfo,FUN=as_tibble,stringsAsFactors=FALSE))
}#end if (! is.finite(idxfuel))
nfuels = nrow(fuelinfo)

# List of soil layer
idxsoi   = match("levgrnd",nc_dlist)
if (is.finite(idxsoi)){
   sois     = nc_conn$dim[[idxsoi]]$vals
   nsois    = nc_conn$dim[[idxsoi]]$len
   soikeys  = sprintf("sl_%3.3f",sois)
   soilabs  = c( paste0("paste(",sois[-nsois],"-",sois[-1],")")
               , paste0("paste(",sois[nsois],"-infinity)")
               )#end dbhlabs
   soilabs  = parse(text=soilabs)
}else{
   sois    = numeric(0L)
   nsois   = 0L
   soikeys = character(0L)
   soilabs = character(0L)
}#end if (is.finite(idxdbh))


#---~---
#retrieve age-class variables
#---~---
nc_byage = nc_vlist[grepl(pattern="_AP$",x=nc_vlist)]
nc_pref  = tolower(gsub(pattern="_AP$",replacement="",x=nc_byage))
nc_keep  = nc_pref %in% fatesvar$vnam & (! duplicated(nc_pref))
no_byage = nc_byage[! nc_keep]
nc_byage = nc_byage[  nc_keep]
nbyage   = length(nc_byage)


# by size class and whether total LAI can be calculated given canopy
# and under-story LAI

is_size   = grepl(pattern="_SZ$",x=nc_vlist) | grepl(pattern="_SZPF$",x=nc_vlist)
nc_bydbh  = nc_vlist[is_size]
nc_pref   = gsub(pattern="_SZ$",replacement="",x=nc_bydbh)
nc_pref   = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
nc_pref   = tolower(nc_pref)
nc_keep   = (nc_pref %in% fatesvar$vnam) & (! duplicated(nc_pref))
no_bydbh  = nc_bydbh[! nc_keep]
nc_bydbh  = unique(nc_bydbh[  nc_keep])

vardbh_last = rep(x=FALSE,times=nfatesvar)
for (v in which(fatesvar$is_upc)){
   nc_vnow      = toupper(fatesvar$vnam[v])
   nc_vnow_scls = paste0(nc_vnow,"_SZ")
   nc_vund_scls = paste0(nc_vnow,"_USTORY_SZ")
   nc_vcan_scls = paste0(nc_vnow,"_CANOPY_SZ")
   nc_vnow_scpf = paste0(nc_vnow,"_SZPF")
   nc_vund_scpf = paste0(nc_vnow,"_USTORY_SZPF")
   nc_vcan_scpf = paste0(nc_vnow,"_CANOPY_SZPF")

   # Check whether this variable can be derived from understorey+canopy (and needs to be).
   if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) ){
      nc_bydbh       = unique(c(nc_bydbh,nc_vnow_scls))
      vardbh_last[v] = TRUE
   }else if ( all(c(nc_vund_scpf,nc_vcan_scpf) %in% nc_bydbh ) && (! nc_vnow_scpf %in% nc_bydbh) ){
      nc_bydbh       = unique(c(nc_bydbh,nc_vnow_scpf))
      vardbh_last[v] = TRUE
   }#end if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) )
}#end for (v in which(fatesvar$is_upc))


nbydbh    = length(nc_bydbh)
n_vardbh_last = sum(vardbh_last)

#by pft
is_pft    = grepl(pattern="_SZPF$",x=nc_vlist) | grepl(pattern="_PF$",x=nc_vlist) | grepl(pattern="_$",x=nc_vlist) #for truncated hydro_cflux
nc_bypft  = nc_vlist[is_pft]
nc_pref   = gsub(pattern="_SZPF$",replacement="",x=nc_bypft )
nc_pref   = gsub(pattern="_PF$",replacement="", x=nc_pref  )
nc_pref   = gsub(pattern="_$",replacement="",x=nc_pref     )
nc_pref   = tolower(nc_pref)
nc_keep   = nc_pref %in% fatesvar$vnam & (! duplicated(nc_pref))
no_bypft  = nc_bypft[! nc_keep]
nc_bypft  = unique(nc_bypft[nc_keep])

varpft_last = rep(x=FALSE,times=nfatesvar)
for (v in which(fatesvar$is_upc)){
   nc_vnow      = toupper(fatesvar$vnam[v])
   nc_vnow_scpf = paste0(nc_vnow,"_SZPF")
   nc_vund_scpf = paste0(nc_vnow,"_USTORY_SZPF")
   nc_vcan_scpf = paste0(nc_vnow,"_CANOPY_SZPF")

   # Check whether this variable can be derived from understorey+canopy (and needs to be).
   if ( all( c(nc_vund_scpf,nc_vcan_scpf) %in% nc_bypft ) && (! nc_vnow_scpf %in% nc_bypft) ){
      nc_bypft       = unique(c(nc_bypft,nc_vnow_scpf))
      varpft_last[v] = TRUE
   }#end if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) )
}#end for (v in which(fatesvar$is_upc))

nbypft    = length(nc_bypft)
n_varpft_last = sum(varpft_last)

# by fuel class
nc_byfuel = nc_vlist[grepl(pattern="_FC$",x=nc_vlist)]
nc_pref   = tolower(gsub(pattern="_FC$",replacement="",x=nc_byfuel))
nc_keep   = nc_pref %in% fatesvar$vnam & (! duplicated(nc_pref))
no_byfuel = nc_byfuel[! nc_keep]
nc_byfuel = nc_byfuel[  nc_keep]
nbyfuel   = length(nc_byfuel)

#---~---
#    Retrieve all "drought deciduous phenology variables
#---~---
nc_pref   = tolower(x=nc_vlist)
nc_keep   = nc_pref %in% dphenvar$vorig
no_dphen  = nc_vlist[! nc_keep]
nc_dphen  = nc_vlist[  nc_keep]

# Tally the total number of drought phenology variables.
ndphen = length(nc_dphen)


# all 1D and 2D variables variables
nc_pref   = tolower(x=nc_vlist)
nc1d_keep = nc_pref %in% hlm1dvar$vnam
no_hlm1d  = nc_vlist[! nc1d_keep]
nc_hlm1d  = nc_vlist[  nc1d_keep]
nc2d_keep = nc_pref %in% hlm2dsoi$vnam
no_hlm2d  = nc_vlist[! nc2d_keep]
nc_hlm2d  = nc_vlist[  nc2d_keep]

# Check whether to append "evapotranspiration"
if (  ( all(c("QSOIL","QVEGT","QVEGE") %in% nc_hlm1d) ) && (! "QEVTR" %in% nc_hlm1d) ){
   nc_hlm1d = unique(c(nc_hlm1d,"QEVTR"))
   etr_last = TRUE
}else{
   etr_last = FALSE
}#end if (  ( all(c("QSOIL","QVEGT","QVEGE") %in% nc_hlm1d) ) && (! "QEVTR" %in% nc_hlm1d) )

# Check whether to append ecosystem respiration
if (  ( all(c("FATES_AUTORESP","FATES_HET_RESP") %in% nc_hlm1d) ) && (! "ER" %in% nc_hlm1d) ){
   nc_hlm1d = unique(c(nc_hlm1d,"ER"))
   er_last  = TRUE
}else{
   er_last  = FALSE
}#end if (  ( all(c("AR","HR") %in% nc_hlm1d) ) && (! "ER" %in% nc_hlm1d) )

# Find number of host land model variables
nhlm1d    = length(nc_hlm1d)
nhlm2d    = length(nc_hlm2d)

#---~---
#Initialize variable matrix
#---~---

#by age class
#byage = list()
#for(c in sequence(ncase)){
#  case     = as.character(case_names[c])
  
#  for(a in sequence(nbyage)){
#  nc_nvnow                   = nc_byage[a]
#  nc_pref                    = tolower(gsub(pattern="_AP$",replacement="",x=nc_nvnow))
#  f                          = match(nc_pref,fatesvar$vnam)
#  f_vnam                     = fatesvar$vnam[f]
#  byage[[case]][[f_vnam]]    = matrix(data=NA_real_,nrow=ntstamp_ens,ncol=nages,dimnames=list(NULL,ageinfo$key)) 
#  } #end of a loop
#} #end of c(in sequence(ncase))


# by size class
#bydbh = list()
#for(c in sequence(ncase)){
#  case     = as.character(case_names[c])
#  for(d in sequence(nbydbh)){
#  nc_nvnow        = nc_bydbh[d]
#  nc_pref         = gsub(pattern="_SZ$",replacement="",x=nc_nvnow)
#  nc_pref         = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
#  nc_pref         = tolower(nc_pref)
#  f               = match(nc_pref,fatesvar$vnam)
#  f_vnam          = fatesvar$vnam[f]
#  bydbh[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens,ncol=ndbhs,dimnames=list(NULL,dbhinfo$key))
#   } # end of d
#}


# by pft

bypft = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  for(p in sequence(nbypft)){
  nc_nvnow        = nc_bypft[p]
  nc_pref         = gsub(pattern="_PF$",replacement="",x=nc_nvnow  )
  nc_pref         = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
  nc_pref         = gsub(pattern="_$",replacement="",x=nc_pref )
  nc_pref         = tolower(nc_pref)
  f               = match(nc_pref,fatesvar$vnam)
  f_vnam          = fatesvar$vnam[f]
  bypft[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens,ncol=npfts,
                           dimnames=list(NULL,pftinfo$key)) 
  } # end of p
}

# Initialize list of variables by fuel class

byfuel = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  
  for(f in sequence(nbyfuel)){
  nc_nvnow                   = nc_byfuel[f]
  nc_pref                    = tolower(gsub(pattern="_FC$",replacement="",x=nc_nvnow))
  f                          = match(nc_pref,fatesvar$vnam)
  f_vnam                     = fatesvar$vnam[f]
  byfuel[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens,ncol=nfuels,dimnames=list(NULL,fuelinfo$key)) 
  } #end of a loop
} #end of c(in sequence(ncase))


# Initialise list of variables by soil layer
dphen = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  for (p in sequence(ndphen)){
   nc_nvnow        = nc_dphen[p]
   nc_pref         = tolower(nc_nvnow)
   f               = match(nc_pref,dphenvar$vorig)
   f_vnam          = dphenvar$vnam[f]
   dphen[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens,
                                    ncol=npfts,dimnames=list(NULL,pftinfo$key))
  }#end for (p in sequence(ndphen))
}



# Initialize 1D variables available at the HLM
hlm1d = as_tibble( matrix( data     = NA_real_
                         , nrow     = ntstamp_ens*ncase
                         , ncol     = nhlm1d+1
                         , dimnames = list(NULL,tolower(c(nc_hlm1d,"CASE")))
                         )
                 )

# Initialize soil water status related variables available at the HLM for each soil layer into hlm2dvar
#hlm2d = list()
#for(c in sequence(ncase)){
#  case     = as.character(case_names[c])
  
#  for(m in sequence(nhlm2dsoi)){
#  nc_nvnow        = nc_hlm2d[m]
#  nc_pref         = tolower(nc_nvnow)
#  h               = match(nc_pref,hlm2dsoi$vnam)
#  h_vnam          = hlm2dsoi$vnam[h]
#  hlm2d[[case]][[h_vnam]] = matrix(data=NA_real_,nrow=ntstamp_ens*ncase,ncol=nsois,dimnames=list(NULL,soikeys))
#  }
#}


# Load soil layers
# do a loop to avoid hard-coding here 
#cat0("   - Load soil information")
#slayer = tibble()

#for (c in sequence(ncase)) {
#case     = as.character(case_names[c])
#slayer_now = tibble( zsoi   = c(unlist(ncvar_get(nc=nc_conn, varid='ZSOI',collapse_degen=TRUE))) 
#                   , dzsoi  = c(unlist(ncvar_get(nc=nc_conn,varid='DZSOI',collapse_degen=TRUE ))) 
#                   , bsw    = c(unlist(ncvar_get(nc=nc_conn,varid='BSW',collapse_degen=TRUE )))  
#                   , hksat  = c(unlist(ncvar_get(nc=nc_conn,varid='HKSAT',collapse_degen=TRUE ))) 
#                   , sucsat = c(unlist(ncvar_get(nc=nc_conn,varid='SUCSAT',collapse_degen=TRUE )))  
#                   , watsat = c(unlist(ncvar_get(nc=nc_conn,varid='WATSAT',collapse_degen=TRUE )))  
#                   , case   = c(rep_along(zsoi, case))
#               )#end tibble
#   slayer = rbind(slayer, slayer_now)
#   rm(slayer_now)
#}


# Load indices
scls   = ncvar_get(nc=nc_conn,varid='fates_scmap_levscpf')
pft    = ncvar_get(nc=nc_conn,varid='fates_pftmap_levscpf')
fuel   = ncvar_get(nc=nc_conn,varid='fates_levfuel')
index_scpf = tibble( scls    
                   , pft    
                   )#end data.table;  
rm(scls,pft)
# Close connection
dummy   = nc_close(nc_conn) 

```






```{r,label="retrieve-values",message=FALSE,results='hide'}
if ("nc_conn" %in% ls()){dummy = nc_close(nc_conn); rm(nc_conn)}

#for(n in sequence(ntstamp_each)){

 # w_month      = month(each_tstamp[w])  
 # w_year       = year(each_tstamp[w])
  
  # Find conversion factors for monthly variables.
  
   cmon.day = days_in_month(ens_tstamp)
   cmon.hr  = day.hr  * cmon.day
   cmon.min = day.min * cmon.day
   cmon.sec = day.sec * cmon.day
  

# Open NetCDF connection and retrieve variable names
for(c in sequence(ncase)){
  case     = case_names[c]
  nc_conn  = nc_open(case_files[c])
  nc_nvars = nc_conn$nvars
  nc_vlist = rep(NA_character_,times=nc_nvars)
  for (v in sequence(nc_nvars)) nc_vlist[v] = nc_conn$var[[v]]$name


# Read variables by age, and assign current values to the matrix.  
#  for (a in sequence(nbyage)){
#    nc_nvnow            = nc_byage[a]
#    nc_pref             = tolower(gsub(pattern="_AP$",replacement="",x=nc_nvnow))
#    f                   = match(nc_pref,fatesvar$vnam)
#    f_vnam              = fatesvar$vnam[f]
#    f_add0              = eval(parse(text=fatesvar$add0[f]))
#    f_mult              = eval(parse(text=fatesvar$mult[f]))
#    nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
   #subset time to only include results from 2000 to 2020
#    nc_dat              = nc_dat[,,,229:480] 
   #reform the 3d array so when convert to matrix, rows are time by ensemble members, columns are age class 
#    ncm_dat             = aperm(nc_dat, c(2,3,4,1)) 
#    dim(ncm_dat)        = c(ntstamp_ens,nages)
#    nc_dat              = f_add0 + f_mult * ncm_dat
#    byage[[case]][[f_vnam]][,]   = nc_dat
#}#end for (a in sequence(nbyage))  

#---~--- 
   #   Read variables by size, and assign current values to the matrix.  In case LAI
   # is in the list and it is the last variable, we calculate it from canopy and under
   # story, after loading all variables
   #---~---

#for (d in sequence(nbydbh-n_vardbh_last)){
#      nc_nvnow            = nc_bydbh[d]
#      nc_pref             = gsub(pattern="_SZPF$",replacement="",x=nc_nvnow)
#      nc_pref             = gsub(pattern="_SZ$",replacement="",x=nc_pref )
#      nc_pref             = tolower(nc_pref)
#      f                   = match(nc_pref,fatesvar$vnam)
#      f_vnam              = fatesvar$vnam[f]
#      f_add0              = eval(parse(text=fatesvar$add0[f]))
#      f_mult              = eval(parse(text=fatesvar$mult[f]))
#      f_aggr              = match.fun(fatesvar$aggr[f])
#      nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow)
#      nc_dat              = nc_dat[,,,229:480] 
#      ncm_dat             = aperm(nc_dat,c(1,2,4,3))
#      dim(ncm_dat)        = c(ntstamp_ens, ndbhs)
#      nc_dat              = f_add0 + f_mult * ncm_dat
#      bydbh[[case]][[f_vnam]][,]  = nc_dat #here we skip aggregating data for size class for by SZPF variable as we                                       #only have 1 PFT, so even SZPF is also a true size class variable
#   }#end for (d in sequence(nbydbh-laidbh_last))

#    Loop through variables to be added last
#   for (v in which(vardbh_last)){
      # Retrieve variable and build understory and canopy variables.
#      v_vnam = fatesvar$vnam[v]
#      v_vund = paste0(v_vnam,"_ustory")
#      v_vcan = paste0(v_vnam,"_canopy")

      # Aggregate data
#      bydbh[[case]][[v_vnam]] = bydbh[[case]][[v_vund]] + bydbh[[case]][[v_vcan]]
#   }#end for (v in which(vardbh_last))
  
 

#---~---
# read variables by PFT
#---~---
  for (p in sequence(nbypft-n_varpft_last)){
      # Load variable information
      nc_nvnow = nc_bypft[p]
      is_szpf  = grepl(pattern="_SZPF$",x=nc_nvnow)
      nc_pref  = tolower(gsub(pattern="_SZPF$",replacement="",x=nc_nvnow))
      nc_pref  = gsub(pattern="_PF$",replacement="",x=nc_pref )
      nc_pref  = gsub(pattern="_$",replacement="",x=nc_pref )
      f        = match(nc_pref,fatesvar$vnam)
      f_vnam   = fatesvar$vnam[f]
      f_add0   = eval(parse(text=fatesvar$add0[f]))
      f_mult   = eval(parse(text=fatesvar$mult[f]))
      nc_dat   = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
      nc_dat   = nc_dat[,,,229:480] 
      if(is_szpf){
      ncm_dat      = aperm(nc_dat,c(1,2,4,3))
      dim(ncm_dat) = c(ntstamp_ens,ndbhs)
      ncm_dat      = f_add0 + f_mult * ncm_dat
      # Aggregate data across size classes
      nc_aggr           = rowSums(ncm_dat, na.rm=TRUE)
      bypft[[case]][[f_vnam]][] = nc_aggr
      }else{
      nc_dat  = as.vector(nc_dat)
      nc_dat  = f_add0 + f_mult * nc_dat
      nc_dat  = matrix(nc_dat,ncol=npfts)
      bypft[[case]][[f_vnam]][] = nc_dat
      }
}#end for (p in sequence(nbypft-n_varpft_last))

# Loop through variables to be added last
  for (v in which(varpft_last)){
      # Retrieve variable and build understory and canopy variables.
    v_vnam = fatesvar$vnam[v]
    v_vund = paste0(v_vnam,"_ustory")
    v_vcan = paste0(v_vnam,"_canopy")

  # Aggregate data
   bypft[[case]][[v_vnam]] = bypft[[case]][[v_vund]] + bypft[[case]][[v_vcan]]
   }#end for (v in which(varpft_last))
  

   # Read variables by fuel class, and assign current values to the matrix.  
# for (f in sequence(nbyfuel)){
#    nc_nvnow            = nc_byfuel[f]
#    nc_pref             = tolower(gsub(pattern="_FC$",replacement="",x=nc_nvnow))
#    f                   = match(nc_pref,fatesvar$vnam)
#    f_vnam              = fatesvar$vnam[f]
#    f_add0              = eval(parse(text=fatesvar$add0[f]))
#    f_mult              = eval(parse(text=fatesvar$mult[f]))
#    nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
   #subset time to only include results from 2000 to 2020
#    nc_dat              = nc_dat[,,,229:480] 
   #reform the 3d array so when convert to matrix, rows are time by ensemble members, columns are fuel class 
#    ncm_dat             = aperm(nc_dat, c(1,2,4,3)) 
#    dim(ncm_dat)        = c(ntstamp_ens, nfuels)
#    ncm_dat              = f_add0 + f_mult * ncm_dat
#    byfuel[[case]][[f_vnam]][,]   = ncm_dat
#}#end for (f in sequence(nbyfuel))  


 # Read drought-deciduous phenology variables (by PFT), and assign current values to the matrix.
   # We use "rev" because the first soil layer is the deepest for the R output.
#   for (p in sequence(ndphen)){
#      nc_nvnow            = nc_dphen[p]
#      nc_pref             = tolower(nc_nvnow)
#      f                   = match(nc_pref,dphenvar$vorig)
#      f_vnam              = dphenvar$vnam[f]
#      f_add0              = eval(parse(text=dphenvar$add0[f]))
#      f_mult              = eval(parse(text=dphenvar$mult[f]))
#      nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
#      dphen[[case]][[f_vnam]][] = f_add0 + f_mult * rev(nc_dat)
      #dphen[[case]][[f_vnam]][,] = f_add0 + f_mult * rev(nc_dat)
#   }#end for (a in sequence(nbyage))


# Read 1D variables
 
  for (v in sequence(nhlm1d-etr_last-er_last)){
  
  nc_nvnow            = nc_hlm1d[v]
  nc_pref             = tolower(x=nc_nvnow)
  h                   = match(nc_pref,hlm1dvar$vnam)
  h_vnam              = hlm1dvar$vnam[h]
  h_add0              = eval(parse(text=hlm1dvar$add0[h]))
  h_mult              = eval(parse(text=hlm1dvar$mult[h]))
  nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
  nc_dat              = nc_dat[,,229:480] 
  ncm_dat             = matrix(nc_dat, ncol=1)
  
  if (c==1){
         hlm1d[[h_vnam]][1:ntstamp_ens]   = h_add0 + h_mult * ncm_dat
         hlm1d$case[1:ntstamp_ens]        = case
     }else{
         start = (c-1)*ntstamp_ens+1
         end   = c*ntstamp_ens
         hlm1d[[h_vnam]][start:end] = h_add0 + h_mult * ncm_dat
         hlm1d$case[start:end]      = case
         }
  
}#for (h in sequence(nhlm1d-etr_last-et_last))
  
 # Find total ET.
   if (etr_last){
      if(c==1){
         hlm1d$qevtr[1:ntstamp_ens] =  hlm1d$qvege[1:ntstamp_ens] + 
                                       hlm1d$qvegt[1:ntstamp_ens] + 
                                       hlm1d$qsoil[1:ntstamp_ens]
      }else{
         start = (c-1)*ntstamp_ens+1
         end   = c*ntstamp_ens
         hlm1d$qevtr[start:end] = hlm1d$qvege[start:end] + 
                                  hlm1d$qvegt[start:end] + 
                                  hlm1d$qsoil[start:end]
      }
   }#end if (etr_last)

   # Find total ET.
   if (er_last){
      if(c==1){
         hlm1d$er[1:ntstamp_ens] =  hlm1d$fates_autoresp[1:ntstamp_ens] + 
                                    hlm1d$fates_het_resp[1:ntstamp_ens]
      }else{
         start = (c-1)*ntstamp_ens+1
         end   = c*ntstamp_ens
         hlm1d$er[start:end] = hlm1d$fates_autoresp[start:end] + 
                               hlm1d$fates_het_resp[start:end] 
               }
   }#end if (er_last)


# Read 2D variables
#for (m in sequence(nhlm2d)){ 
#  nc_nvnow            = nc_hlm2d[m]
#  nc_pref             = tolower(x=nc_nvnow)
#  h                   = match(nc_pref,hlm2dsoi$vnam)
#  h_vnam              = hlm2dsoi$vnam[h]
#  h_add0              = eval(parse(text=hlm2dsoi$add0[h]))
#  h_mult              = eval(parse(text=hlm2dsoi$mult[h]))
#  nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
#  nc_dat              = nc_dat[,181:432,] 
#  ncm_dat             = aperm(nc_dat, c(2,3,1))
#  dim(ncm_dat)        = c(ntstamp_each*ninst, nsois)
#  nc_dat              = h_add0 + h_mult * ncm_dat
#  hlm2d[[case]][[h_vnam]][,]  = nc_dat
#}#end of for (m in sequence(nhlm2d))
# Close connection
dummy   = nc_close(nc_conn)

  
} #end of for (c in sequence(ncase))

## add time for hlm1d
hlm1d$time = rep(ens_tstamp,times=ncase)
hlm1d$lon  = rep(lon_vec, times=nstep*ncase)
hlm1d$lat  = rep(lat_vec, times=nstep*ncase)
fwrite(hlm1d,file.path(region_path,paste0("region_hlm1d.csv")))

```






```{r,label="scale-age",message=FALSE,results='hide'}
cat0(" + Rescale variables by age class.")

for(c in sequence(ncase)){
  case = case_names[c]
  for (a in sequence(nbyage)){
#--- Match variables
  f        = match(names(byage[[case]])[a],fatesvar$vnam)
  f_vnam   = fatesvar$vnam [f]
  f_desc   = fatesvar$desc [f]
  f_stack  = fatesvar$stack[f]
   #---~---

   #--- Proceed only if the variable is stacked and not the patch area.
  if (f_stack && (! (f_vnam %in% "fates_patcharea"))){
      cat0("   - ",f_desc,".")
      byage[[case]][[f_vnam]] = byage[[case]][[f_vnam]] * byage[[case]]$fates_patcharea
  }#end if (f_stack && (! (f_vnam %in% "patch_area")))
}#end for (a in sequence(nbyage))
  
}



```







```{r,label="melt-data",message=FALSE,results='hide'}
#cat0(" + Turn age-dependent matrices into tibble objects.")
#age_melt  = NULL

#for (c in sequence(ncase)){
#  case = case_names[c]
#  now_agemelt = NULL
#  for (a in sequence(nbyage)){
   #--- Match variables.
#  f        = match(names(byage[[case]])[a],fatesvar$vnam)
#  f_vnam   = fatesvar$vnam[f]
#  f_desc   = fatesvar$desc[f]
#  cat0("   - ",f_desc,".")
      
#--- Create molten data table for this variable.  
#  now_age           = as_tibble(byage[[case]][[f_vnam]])
#  now_age$time      = rep(each_tstamp, time=ninst)
#  now_age$case      = rep_along(time, x = case)
#  now_age$ens_label = rep(ens_labels, each = ntstamp_each)
#  now_melt          = as_tibble(reshape2::melt(data=now_age,id.vars=c("case","ens_label",       
#                              "time"),variable.name="age",value.name=f_vnam))
      #--- Merge data table
  
#  if (is.null(now_agemelt)){
#    now_agemelt = now_melt
#  }else{
#        now_agemelt = as_tibble(merge(x=now_agemelt,y=now_melt,by=c("case","ens_label","time","age"),all=TRUE))
#  }#end if (is.null(age_melt))
  
#  }#end for (a in sequence(nbyage))
#age_melt = rbind(age_melt,now_agemelt)
#}



#byage = age_melt 
#rm(age_melt)
#invisible(gc())

#cat0(" + Turn size-dependent matrices into data tables.")
#dbh_melt = NULL
#for (c in sequence(ncase)){
#  case = case_names[c]
#for (d in sequence(nbydbh)){
#  f        = match(names(bydbh[[case]])[d],fatesvar$vnam)
#  f_vnam   = fatesvar$vnam[f]
#  f_desc   = fatesvar$desc[f]
#  cat0("   - ",f_desc,".")
   
#  now_dbh           = as_tibble(bydbh[[case]][[d]])
#  now_dbh$time      = rep(each_tstamp, time=ninst)
#  now_dbh$case      = rep_along(time, x = case)
#  now_dbh$ens_label = rep(ens_labels, each = ntstamp_each)
#  now_melt          = as_tibble(reshape2::melt(data=now_dbh,id.vars=c("case", "ens_label",  "time"),variable.name="dbh",value.name=f_vnam))
  
#  if (is.null(dbh_melt)){
#      dbh_melt = now_melt
#      }else{
#         dbh_melt = merge(x=dbh_melt,y=now_melt,by=c("case", "ens_label","time","dbh"),all=TRUE)
#      }#end if (is.null(scls.melt))
#}#end for (d in sequence(nbydbh))

#}

#bydbh = dbh_melt
#rm(dbh_melt)

cat0(" + Turn PFT-dependent matrices into data tables.")
pft_melt = NULL
for (c in sequence(ncase)){
  case = case_names[c]
  now_pftmelt = NULL
  for (p in sequence(nbypft)){
   
  f        = match(names(bypft[[case]])[p],fatesvar$vnam)
  f_vnam   = fatesvar$vnam[f]
  f_desc   = fatesvar$desc[f]
  cat0("   - ",f_desc,".")
  
  now_pft           = as_tibble(bypft[[case]][[p]])
  now_pft$lon       = rep(lon_vec,times=nstep)
  now_pft$lat       = rep(lat_vec,times=nstep)
  now_pft$time      = ens_tstamp
  now_pft$case      = rep_along(time, x=case)
  now_melt          = as_tibble(reshape2::melt(data=now_pft,id.vars=c("case","lon","lat","time"),variable.name="pft",value.name=f_vnam))

  if (is.null(now_pftmelt)){
      now_pftmelt = now_melt
      }else{
         now_pftmelt = as_tibble(merge(x=now_pftmelt,y=now_melt,by=c("case", "lon", "lat","time", "pft"),all=TRUE))
      }
}#end for (p in sequence(nbypft))
  pft_melt = rbind(pft_melt,now_pftmelt)
}


bypft = pft_melt
rm(pft_melt)
invisible(gc())

#cat0(" + Turn fuel-class matrices into tibble objects.")
#fuel_melt  = NULL

#for (c in sequence(ncase)){
#  case = case_names[c]
#  now_fuelmelt = NULL
#  for (u in sequence(nbyfuel)){
   #--- Match variables.
#  f        = match(names(byfuel[[case]])[u],fatesvar$vnam)
#  f_vnam   = fatesvar$vnam[f]
#  f_desc   = fatesvar$desc[f]
#  cat0("   - ",f_desc,".")
      
#--- Create molten data table for this variable.  
#  now_fuel            = as_tibble(byfuel[[case]][[f_vnam]])
#  now_fuel$lon        = rep(lon_vec,time=nstep)
#  now_fuel$lat        = rep(lat_vec,time=nstep)
#  now_fuel$time       = ens_tstamp
#  now_fuel$case       = rep_along(time, x = case)
#  now_melt            = as_tibble(reshape2::melt(data=now_fuel,id.vars=c("case","lon",       
#                              "lat","time"),variable.name="fuel",value.name=f_vnam))
      #--- Merge data table
  
#  if (is.null(now_fuelmelt)){
#    now_fuelmelt = now_melt
#  }else{
#        now_fuelmelt = as_tibble(merge(x=now_fuelmelt,y=now_melt,by=c("case","lon","lat","time","fuel"),all=TRUE))
#  }#end if (is.null(age_melt))
  
 # }#end for (a in sequence(nbyage))
#fuel_melt = rbind(fuel_melt,now_fuelmelt)
#}



#byfuel = fuel_melt 
#rm(fuel_melt)
#invisible(gc())

# Turn drought-deciduous phenology variables into data tables
#if (! is_tibble(dphen)){
#   cat0(" + Turn drought-deciduous phenology matrices into data tables.")
#   dph_melt = NULL
#   for (c in sequence(ncase)){
#     case = case_names[c]
#     now_dphmelt = NULL
#     for (p in sequence(ndphen)){
      # Match variables.
#      f        = match(names(dphen[[case]])[p],dphenvar$vnam)
#      f_vnam   = dphenvar$vnam[f]
#      f_desc   = dphenvar$desc[f]
#      cat0("   - ",f_desc,".")

      # Create molten data table for this variable.  
#      now_dph           = as_tibble(dphen[[case]][[p]])
#      now_dph$lon       = lon_vec
#      now_dph$lat       = lat_vec
#      now_dph$time      = ens_tstamp
#      now_dph$case      = rep_along(time, x=case)
#      now_melt          = as_tibble(melt(data=now_dph,id.vars=c("case","lon","lat","time"),
#                                    variable.name="pft",value.name=f_vnam))

      # Merge data table
#      if (is.null(now_dphmelt)){
#         now_dphmelt = now_melt
#      }else{
#         now_dphmelt = as_tibble(merge(x=now_dphmelt,y=now_melt,by=c("case","lon","lat","time","pft"),all=TRUE))
#      }#end if (is.null(dph_melt))
#   }#end for (p in sequence(ndphen))
#    dph_melt = rbind(dph_melt,now_dphmelt)
#}# end of for (c insequence(ncase))
   
   # Replace dphen with the tibble object
#   dphen = dph_melt
#   rm(dph_melt)
#   invisible(gc())
#}#end if (! is_tibble(dphen))

#cat0(" + Turn soil layer-dependent matrices into data tables.")
#soi_melt = NULL
#for (c in sequence(ncase)){
#  case = case_names[c]
#  now_soimelt = NULL
#  for (s in sequence(nhlm2d)){
#  h        = match(names(hlm2d[[case]])[s],hlm2dsoi$vnam)
#  h_vnam   = hlm2dsoi$vnam[h]
#  h_desc   = hlm2dsoi$desc[h]
#  cat0("   - ",h_desc,".")
   
#  now_soi           = as_tibble(hlm2d[[case]][[s]])
#  now_soi$lon       = lon_vec
#  now_soi$lat       = lat_vec
#  now_soi$time      = ens_tstamp
#  now_soi$case      = rep_along(time, x=case)
#  now_melt     = as_tibble(reshape2::melt(data=now_soi,id.vars=c("case", "lon",                 
#                                             "lat","time"),variable.name="soi",value.name=h_vnam))
#  if (is.null(now_soimelt)){
#      now_soimelt = now_melt
#      }else{
#         now_soimelt = merge(x=now_soimelt,y=now_melt,by=c("case", "lon","lat","time","soi"),all=TRUE)
#      }#end if (is.null(scls.melt))
#  }#end for (s in sequence(nsoi))
#  soi_melt = rbind(soi_melt,now_soimelt)
  
#}


#bysoi = soi_melt
#rm(soi_melt)
#invisible(gc())


cat0(" + Convert classes to integers.")
#byage$age = as.integer(byage$age)
#bydbh$dbh = as.integer(bydbh$dbh)
bypft$pft   = as.integer(bypft$pft)
#byfuel$fuel = as.integer(byfuel$fuel)
#dphen$pft = as.integer(dphen$pft)
#bysoi$soi = as.integer(bysoi$soi)

## write out tables so can use later
#fwrite(bypft,file.path(region_path,paste0("region_bypft.csv")))
#fwrite(byfuel,file.path(region_path,paste0("region_byfuel.csv")))

bypft  = fread(file.path(region_path,"region_bypft.csv"))
hlm1d  = fread(file.path(region_path,"region_hlm1d.csv"))
#byfuel = fread(file.path(region_path,"region_byfuel.csv"))

```









```{r,label="load-observations",message=FALSE,results='hide'}
cat0(" + Load regional data from ",obs_main,".")


## load FRAP regional burned fraction range

cags_bf    = fread(frap_grass)       
cags_bf    = cags_bf                       %>% 
             filter(year %in% simu_time)   %>% 
             mutate(bfrac=bfrac
                   ,lon=round(lon,digits=5)
                   ,lat=round(lat,digits=5)) 
cags_bf$lon_f = factor(cags_bf$lon,levels=unique(cags_bf$lon))
cags_bf$lat_f = factor(cags_bf$lat,levels=unique(cags_bf$lat))

             
cags_dmean = cags_bf                                   %>%
             group_by(lon_f,lat_f)                     %>% 
             mutate(bfrac=mean(bfrac,na.rm=TRUE))      %>% 
             ungroup()                                  
             

cags_qupr  = cags_bf                                   %>% 
             group_by(lon_f,lat_f)                     %>% 
             summarize(bfrac_qupr=quantile(bfrac
                                          ,probs=qupr_ribbon
                                          ,na.rm=TRUE)) %>% 
             ungroup()                                 

cags_qlwr  = cags_bf                                   %>% 
             group_by(lon_f,lat_f)                     %>% 
             summarize(bfrac_qlwr=quantile(bfrac
                                          ,probs=qlwr_ribbon
                                          ,na.rm=TRUE)) %>% 
             ungroup()

cags_dmean = cags_dmean                                  %>% 
             left_join(cags_qupr,by=c("lon_f","lat_f"))  %>% 
             left_join(cags_qlwr,by=c("lon_f","lat_f"))  %>% 
             dplyr::select(-c(lon_f,lat_f,year,mask))    %>% 
             distinct()




## range of annual burned fraction by each year using regional data         
bf_quantyr = cags_bf                                 %>% 
             group_by(year)                          %>%
             summarize(yr_qupr = quantile(bfrac,probs=qupr_ribbon,na.rm=TRUE)
                      ,yr_qlwr = quantile(bfrac,probs=qlwr_ribbon,na.rm=TRUE)
                      ,yr_mean = mean(bfrac,na.rm=TRUE)) %>% 
             ungroup()


## load SIF re-sampled regional GPP




## load MODIS re-sampled regional LAI 




## join burned fraction, GPP, and LAI

```






```{r,label='load-biomass-data-nultiple-sites',message=FALSE,results='hide'}


## load pepper wood preserve data
agb = fread(agbP_path)
agb = agb                                 %>% 
      mutate(date       = ymd(date)
            ,year       = year(date)
            ,month      = month(date)
            ,biomass    = biomass/2
            ,biomass_sd = biomass_sd/2)   %>% 
      dplyr::select(-date)                %>% 
      group_by(year,month)                %>% 
      summarise_all(mean,na.rm=TRUE)      %>% 
      ungroup()

## load seabloom et al. data
seablm_data = fread(agbS_path)
sierra_amean = seablm_data                                   %>% 
              filter(site == "sier.us" & trt == "Control")   %>% 
              mutate(live = (10^live_lg)*g2kg/2)             %>% 
              group_by(year)                                 %>% 
              summarize(agbs      = mean(live,na.rm=TRUE)
                       ,agbs_qupr = quantile(live,probs=qupr_ribbon
                                            ,na.rm=TRUE)
                       ,agbs_qlwr = quantile(live,probs=qlwr_ribbon
                                            ,na.rm=TRUE))

## maybe instead comparing AGB we compare how AGB is related to precipitation and temperature? 
              

```







```{r,label="fates-variable-mean",fig.height=50, fig.width=40,message=FALSE}

mortout    = names(bypft)[grep(pattern="mortality",x=names(bypft))]
find_mort = function(x,n){
   mort_mon = ifelse( test = n == 0, yes = 0., no = pmin(1.,x/n/12.))
   mort_year = 100. * (1. - (1. - mort_mon)^12)
   return(mort_year)
}#end function find_mort


all    =     bypft                                            %>%  
             mutate(month = month(time)
                   ,year  = year(time))                       %>% 
             mutate_at(all_of(mortout)
                      ,~ find_mort(x=.x,n=.data$fates_nplant)) %>% 
             mutate(lon = round(lon,5)
                   ,lat = round(lat,5))
  
all$lon_f = factor(all$lon,levels=unique(all$lon))
all$lat_f = factor(all$lat,levels=unique(all$lat))

orig_xy   = hlm1d  %>% dplyr::select(lon,lat) %>% distinct() %>% 
            mutate(lon=round(lon,5)
                  ,lat=round(lat,5)
                  ,lon_f = factor(lon,levels=unique(lon))
                  ,lat_f = factor(lat,levels=unique(lat)))

cell_mmean = all                                                     %>% 
            group_by(case,lon_f,lat_f,pft,month)                     %>%  
  summarize( fates_ddbh_canopy            = mean(fates_ddbh_canopy)
           , fates_ddbh_ustory            = mean(fates_ddbh_ustory)
           , fates_gpp                    = mean(fates_gpp)
           , fates_lai_canopy             = mean(fates_lai_canopy)
           , fates_lai_ustory             = mean(fates_lai_ustory)
           , fates_mortality_background   = mean(fates_mortality_background)
           , fates_mortality_cstarv       = mean(fates_mortality_cstarv)
           , fates_mortality_hydraulic    = mean(fates_mortality_hydraulic)
           , fates_mortality_termination  = mean(fates_mortality_termination)
           , fates_mortality_canopy       = mean(fates_mortality_canopy)
           , fates_mortality_ustory       = mean(fates_mortality_ustory)
           , fates_npp                    = mean(fates_npp)
           , fates_vegc_aboveground       = mean(fates_vegc_aboveground)
           , fates_lai                    = mean(fates_lai))  %>% 
  ungroup()                                                                   

cell_mmean = orig_xy %>% left_join(cell_mmean,by=c("lon_f","lat_f")) 


 cell_amean = all                                                        %>% 
              group_by(case,lon_f,lat_f,pft,year)                        %>% 
              summarize(fates_ddbh_canopy           = mean(fates_ddbh_canopy)
                       ,fates_ddbh_ustory           = mean(fates_ddbh_ustory)
                       ,fates_gpp                   = mean(fates_gpp)
                       ,fates_lai_canopy            = mean(fates_lai_canopy)
                       ,fates_lai_ustory            = mean(fates_lai_ustory)
                       ,fates_vegc_aboveground      = mean(fates_vegc_aboveground)
                       ,fates_lai                   = mean(fates_lai))   %>%
               ungroup()                                                 
 cell_amean = orig_xy %>% left_join(cell_amean,by=c("lon_f","lat_f"))
 
 
 cell_mean  = all                                                         %>% 
              group_by(case,lon_f,lat_f,pft)                              %>% 
              summarize(fates_gpp                   = mean(fates_gpp)
                       ,fates_lai_canopy            = mean(fates_lai_canopy)
                       ,fates_lai_ustory            = mean(fates_lai_ustory)
                       ,fates_vegc_aboveground      = mean(fates_vegc_aboveground)
                       ,fates_lai                   = mean(fates_lai))   %>%
               ungroup()                                                 
 cell_mean   = orig_xy  %>% left_join(cell_mean,by=c("lon_f","lat_f"))
              



  
```











```{r,label="hlm-variable-mean",results='hide'}

hlm_var   = c("fates_gpp","fates_vegc_aboveground",
              "eflx_lh_tot","elai","fsh")
fire_month = c("6","7","8","9","10")
growing_sea   = c("3","4","5")

hlm1d$lon_f = factor(hlm1d$lon,levels=unique(hlm1d$lon))
hlm1d$lat_f = factor(hlm1d$lat,levels=unique(hlm1d$lat))

bf_mod   =  hlm1d                                                                  %>% 
            dplyr::select(all_of(c("case","time","lon","lat",
                                   "lon_f","lat_f","fates_burnfrac")))             %>% 
            mutate(month=month(time))                                              %>% 
            filter(month %in% fire_month)                                          %>% 
            dplyr::select(-month)


hlm_scy    = hlm1d                                                                %>% 
             dplyr::select(all_of(c("case","time","lon","lat",
                                    "lon_f","lat_f", hlm_var)))                   %>% 
             mutate(month=month(time))                                            %>% 
             filter(month%in%growing_sea)                                         %>% 
             dplyr::select(-month)



mean_bf =   bf_mod                                     %>%
            group_by(case,lon_f,lat_f)                 %>%
            dplyr::select(-time)                       %>%
            mutate(fates_burnfrac=mean(fates_burnfrac)
                  )                                    %>% 
            ungroup()                                  %>% 
            dplyr::select(-c(lon_f,lat_f))             %>% 
            distinct()
mean_bfall = hlm1d                                       %>% 
             group_by(case,lon_f,lat_f)                  %>% 
             mutate(fates_burnfrac=mean(fates_burnfrac)) %>% 
             ungroup()                                   %>% 
             dplyr::select(case,lon,lat,fates_burnfrac)  %>% 
             distinct()                               

sum_bfall  = hlm1d                                     %>% 
             group_by(case,lon_f,lat_f)                %>% 
             mutate(bfrac=sum(fates_burnfrac))         %>% 
             ungroup()                                 %>% 
             dplyr::select(case,lon,lat,bfrac)         %>% 
             distinct()                                %>% 
             mutate(bfrac=bfrac*yr.day) 


mean_hlm   = hlm_scy                          %>% 
             group_by(case,lon_f,lat_f)       %>% 
             dplyr::select(-time)             %>% 
             mutate(fates_gpp     = mean(fates_gpp)
                   ,fates_vegc_aboveground = mean(fates_vegc_aboveground)
                   ,eflx_lh_tot = mean(eflx_lh_tot)
                   ,elai = mean(elai)
                   ,fsh = mean(fsh))          %>% 
             distinct()

allmean_hlm = hlm1d                                      %>% 
              dplyr::select(all_of(c("case","lon","lat",
                          "lon_f","lat_f",hlm_var)))     %>% 
              group_by(case,lon_f,lat_f)                 %>% 
              mutate(fates_gpp     = mean(fates_gpp)
                   ,fates_vegc_aboveground = mean(fates_vegc_aboveground)
                   ,eflx_lh_tot = mean(eflx_lh_tot)
                   ,elai = mean(elai)
                   ,fsh = mean(fsh))                     %>% 
               distinct()


## mean of each variable from either peak growing season or fire season
mean_pseason = mean_hlm %>% left_join(mean_bf,by=c("case","lon","lat")) %>% ungroup()

## mean of each variable from all months
mean_allseson = allmean_hlm  %>% left_join(mean_bfall,by=c("case","lon","lat")) %>% ungroup()


```













```{r,label="plot-HLM-mean"}
cat0(" + Plot  mean of HLM variables.")

#leg_colours = c(site_colour,ens_colour)
#leg_labels  = c("Site",case_desc)
ca_co    = USAboundaries::us_counties(resolution = "high", states = "CA")
wgs = "+init=EPSG:4326"
emean_loop = which(hlm1dvar$vnam %in% c(hlm_var,"fates_burnfrac"))


gg_emean   = list()
for(c in sequence(ncase)){
  case_now = case_names[c]
  desc_now = case_desc[c]
mmean_now        = mean_allseson %>% filter(case == case_now) 
for (h in emean_loop){
   h_vnam   = hlm1dvar$vnam[h] 
   h_desc   = hlm1dvar$desc [h]
   h_short  = hlm1dvar$short[h]
   h_unit   = hlm1dvar$unit [h]
   h_vqlwr  = paste0(h_vnam,"_qlwr")
   h_vqupr  = paste0(h_vnam,"_qupr")
   h_legend =  1
   cat0("   - ",h_desc,".")
   var_df   = mmean_now %>% dplyr::select(all_of(c("lon","lat",h_vnam)))
   
   
   fil_var  = matrix(var_df[[h_vnam]], nrow=147,ncol=151)
   x_arr    = matrix(var_df$lon,nrow=147,ncol=151)
   y_arr    = matrix(var_df$lat,nrow=147,ncol=151)
   wrf_star = st_as_stars(fil_var)
   wrf_star = st_as_stars(wrf_star, curvilinear=list(X1=x_arr,X2=y_arr), crs=wgs)
   wrf_sf   = st_as_sf(wrf_star,as_points=FALSE,na.rm=FALSE)
   
   gg_now = ggplot()
   gg_now = gg_now + geom_sf(data=wrf_sf,aes(fill=A1),color=NA)
   gg_now = gg_now + coord_sf(crs=st_crs(wgs)) 
   gg_now = gg_now + scale_fill_continuous(low="blue", high="red", guide="colorbar",na.value="white")
   gg_now = gg_now + geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) 
   gg_now = gg_now + labs(title=desc_now
                         , x="Longitude"
                         , y="Latitude"
                         , fill=desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right" 
                          , panel.background  = element_blank()
                          , panel.border      = element_rect(linewidth = 1.6, fill=NA)
                          , axis.text.x       = element_text( size   = gg_ptsz-5
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                          , axis.text.y       = element_text( size   = gg_ptsz-5
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                          , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(case_now,h_vnam,"allmean.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot     = gg_now
                    , device   = gg_device[d]
                    , path     = secy_path
                    , width    = gg_widthn*0.6
                    , height   = gg_height*0.6
                    , units    = gg_units
                    , dpi      = gg_depth
                    )
   }
  gg_emean[[h_vnam]] = gg_now      

}
}


# If sought, plot images on screen
if (gg_screen) gg_emean

```







```{r,label="relative-bf"}

bf_simu = mean_hlm %>% dplyr::select(case,lon,lat,fates_burnfrac) %>% 
          mutate(lon=round(lon,digits=5)
                ,lat=round(lat,digits=5))     
bfsum_simu = sum_bfall %>% dplyr::select(case,lon,lat,bfrac) %>% 
             mutate(lon=round(lon,digits=5)
                   ,lat=round(lat,digits=5))

bfsum_ob   = cags_bf %>% group_by(lon_f,lat_f)   %>% 
             mutate(bfrac_ob=sum(bfrac,na.rm=TRUE)) %>% 
             ungroup()                           %>% 
             dplyr::select(lon,lat,bfrac_ob)     %>% 
             distinct()                          %>% 
             mutate(bfrac_ob=bfrac_ob*yr.day)

bf_comm = bf_simu  %>% left_join(cags_dmean,by=c("lon","lat")) %>% 
          mutate(rela_bf = fates_burnfrac/bfrac)               %>% 
          mutate(rela_bf = ifelse(is.infinite(rela_bf),NA,rela_bf))

bfall_comm = bfsum_simu %>% left_join(bfsum_ob,by=c("lon","lat"))
 
for(n in sequence(ncase)){
  case_now = case_names[n]
  bf_now = bf_comm %>% filter(case==case_now)
 # bf_noext = bf_comm %>% mutate(rela_bf=ifelse(rela_bf>100 | rela_bf<0.01,NA,rela_bf))


 fil_var  = matrix(bf_now$rela_bf, nrow=147,ncol=151)
 x_arr    = matrix(bf_now$lon,nrow=147,ncol=151)
 y_arr    = matrix(bf_now$lat,nrow=147,ncol=151)
 wrf_star = st_as_stars(fil_var)
 wrf_star = st_as_stars(wrf_star, curvilinear=list(X1=x_arr,X2=y_arr), crs=wgs)
 wrf_sf   = st_as_sf(wrf_star,as_points=FALSE,na.rm=FALSE)
 ca_co    = USAboundaries::us_counties(resolution = "high", states = "CA")


 p = ggplot() + geom_sf(data=wrf_sf,colour="grey50", aes(fill=A1),lwd=0)+
  coord_sf(crs=st_crs(wgs)) + 
  scale_fill_continuous(low="blue", high="red", 
                                  guide="colorbar",na.value="white")+
  geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) +
  labs(x="",y="") +
  guides(fill=guide_legend(title="Relative burned fraction")) +
  theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
  theme( legend.position   = "right" 
       , panel.background = element_rect(size = 1.6, fill = NA)
       , panel.border = element_rect(size = 1.6, fill=NA)
       , axis.text.x   = element_text( size   = gg_ptsz
       , margin = unit(rep(0.35,times=4),"cm"))#end element_text
       , axis.text.y       = element_text( size   = gg_ptsz
       , margin = unit(rep(0.35,times=4),"cm")
      )#end element_text
      , axis.ticks.length = unit(-0.25,"cm")
      ) #end theme
   p
   fname = paste0(case_now,"-relativeBF.png")
   ggsave(p, filename=fname,path=secy_path,width=gg_widthn,height=gg_height,dpi=gg_depth,device=gg_device[1])
}

## seems case 1 is the best given relative BF
both    = bf_comm %>% filter(case==3) %>% mutate(diff = (fates_burnfrac-bfrac)/bfrac)
both_nox= both    %>% mutate(rela_bf=ifelse(rela_bf>100 | rela_bf< 0.01,NA,rela_bf))
frap    = bf_comm %>% filter(case==3) %>% dplyr::select(lon,lat,bfrac) %>% mutate(bfrac = bfrac*yr.day)
model   = bf_comm %>% filter(case==3) %>% dplyr::select(lon,lat,fates_burnfrac) %>% mutate(bfrac = fates_burnfrac*yr.day)

 model_var  = matrix(model$bfrac, nrow=147,ncol=151) 
 x_arr    = matrix(model$lon,nrow=147,ncol=151)
 y_arr    = matrix(model$lat,nrow=147,ncol=151)
 model    = st_as_stars(model_var)
 model    = st_as_stars(model, curvilinear=list(X1=x_arr,X2=y_arr), crs=wgs)
 model_sf   = st_as_sf(model,as_points=FALSE,na.rm=FALSE)
 ca_co    = USAboundaries::us_counties(resolution = "high", states = "CA")

par(mar=c(0,0,0,0))
p1 = ggplot() + geom_sf(data=model_sf,colour="grey50", aes(fill=A1),lwd=0)+
  scale_fill_continuous(low="blue", high="red", 
                        guide="colorbar",na.value="white")+
  geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) +
  labs(x="",y="",title=expression("FATES burned fraction"~"("*year^-1*")")) +
  guides(fill=guide_legend(title="")) +
  theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
  theme( 
        legend.position   = "bottom" 
       , panel.background = element_rect(fill=NA)
       , panel.border = element_rect(fill=NA)
       , axis.text.x   = element_text( size   = gg_ptsz-5
       , margin = unit(rep(0.35,times=4),"cm"))#end element_text
       , axis.text.y       = element_text( size   = gg_ptsz-5
       , margin = unit(rep(0.35,times=4),"cm")
      )#end element_text
      , axis.ticks.length = unit(-0.25,"cm")
      ) #end theme
 p1 
fname="medianBF-bf.png"
ggsave(p1, filename=fname,path=secy_path,width=gg_widthn*0.6,height=gg_height*0.6,dpi=gg_depth,device=gg_device[1])

par(mar=c(0,0,0,0))
p2=ggplot() + geom_sf(data=model_sf,colour="grey50", aes(fill=A1),lwd=0)+
  scale_fill_continuous(low="blue", high="red", 
                        guide="colorbar",na.value="white")+
  geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) +
  labs(x="",y="")  +
  guides(fill=guide_legend(title="")) +
  theme_grey( base_size = gg_ptsz-10, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
  theme( legend.position   = c(0.6,0.7)
       , legend.key.size = unit(0.1, "cm")
       , panel.background = element_blank()
       , panel.border = element_blank()
       , axis.line=element_blank(),axis.text.x=element_blank()
       , axis.text.y=element_blank(),axis.ticks=element_blank()
       , axis.title.x=element_blank()
       , axis.title.y=element_blank()
      ) #end theme
   p2
   fname="FATES-2000-2020SUM.png"
 ggsave(p2, filename=fname,path=secy_path,width=gg_widthn*0.4,height=gg_height*0.4,dpi=gg_depth,device=gg_device[1])
## FRAP BF plot
 
 fil_var  = matrix(both_nox$bfrac, nrow=147,ncol=151)
 x_arr    = matrix(both_nox$lon,nrow=147,ncol=151)
 y_arr    = matrix(both_nox$lat,nrow=147,ncol=151)
 wrf_star = st_as_stars(fil_var)
 wrf_star = st_as_stars(wrf_star, curvilinear=list(X1=x_arr,X2=y_arr), crs=wgs)
 wrf_sf   = st_as_sf(wrf_star,as_points=FALSE,na.rm=FALSE)
 
par(mar=c(0,4,4,4))
 p3=ggplot() + geom_sf(data=wrf_sf,colour="grey50", aes(fill=A1),lwd=0)+
  scale_fill_continuous(low="blue", high="red", 
                        guide="colorbar",na.value="white")+
  geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) +
  labs(x="",y="")  +
  guides(fill=guide_legend(title="")) +
  theme_grey( base_size = gg_ptsz-10, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
  theme( legend.position   = c(0.6,0.7)
       , legend.key.size = unit(0.1, "cm")
       , panel.background = element_blank()
       , panel.border = element_blank()
       , axis.line=element_blank(),axis.text.x=element_blank()
       , axis.text.y=element_blank(),axis.ticks=element_blank()
       , axis.title.x=element_blank()
       , axis.title.y=element_blank()
      ) #end theme
   
 p3

   p = p2 + p3 + guide_area() + plot_layout(ncol=2,widths = c(4, -1,4))                                               
  
p
   fname = "frap_2000-2020sum.png"
   ggsave(p3, filename=fname,path=secy_path,width=gg_widthn*0.4,height=gg_height*0.4,dpi=gg_depth,device=gg_device[1])

### annual total burned fraction compare to FRAP
mo_amean =  hlm_scy                                                        %>%
            mutate(year=year(time))                                        %>%
            group_by(case,year)                                            %>%
            summarize(bf_amean=sum(fates_burnfrac,na.rm=TRUE))             %>% 
            ungroup()                                                      %>% 
            mutate(bf_amean = bf_amean*yr.day)                             %>% 
            filter(case==3)                                                %>% 
            mutate(case="FATES")
mo_allamean = hlm1d                                                        %>% 
              mutate(year=year(time))                                      %>% 
              group_by(case,year)                                          %>% 
              summarize(bf_amean=sum(fates_burnfrac,na.rm=TRUE))           %>% 
              mutate(bf_amean = bf_amean*yr.day)                           %>%
              ungroup()                                                    %>% 
              mutate(case=case_desc[case])
ob_amean = cags_bf                                                         %>% 
           group_by(year)                                                  %>% 
           summarize(bf_amean=sum(bfrac,na.rm=TRUE))                       %>% 
           mutate(bf_amean=bf_amean*yr.day
                 ,case="FRAP")

amean_comm = rbind(mo_allamean,ob_amean)
#amean_comm$case=factor(amean_comm$case,levels=unique(amean_comm$case))
p = ggplot(amean_comm,aes(year,bf_amean,colour=case)) + geom_line() +
scale_color_manual(name="",values=ens_colour) +
  labs(x="",y="Summed burned fraction")+
  theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)+
  theme( legend.position   = "right" 
       , legend.text = element_text(size=gg_ptsz-5)
       , panel.background = element_rect(size = 1.6, fill = NA)
       , panel.border = element_rect(size = 1.6, fill=NA)
       , axis.text.x   = element_text( size   = gg_ptsz
       , margin = unit(rep(0.35,times=4),"cm"))#end element_text
       , axis.text.y       = element_text( size   = gg_ptsz
       , margin = unit(rep(0.35,times=4),"cm")
      )#end element_text
      , axis.ticks.length = unit(-0.25,"cm")
      ) #end theme
out_name = "summed-BF-all.png"
ggsave(p, filename=out_name,path=secy_path,width=gg_widthn*0.6,height=gg_height*0.6,dpi=gg_depth,device=gg_device[1])

```






```{r,label="plot-mean-fire-variables"}

fire_vars = c("fates_burnfrac","fates_fdi","fates_ros",
              "fates_fuel_amount","fates_fuel_bulkd")

fire_mmean = hlm1d                                                 %>% 
             select(all_of(c("case","ens_label","time",fire_vars)))%>% 
             mutate(case_ens = paste0(case,"_",ens_label)
                   , month = month(time))                          %>% 
             group_by(case,ens_label,case_ens,month)               %>% 
             dplyr::select(-time)                                  %>% 
             summarize_all(mean,na.rm=TRUE)                        %>% 
             ungroup()

fuelmmean_bysize = byfuel                                          %>% 
              mutate(case_ens = paste0(case,"_",ens_label)
                   , month = month(time))                          %>% 
              group_by(case,month,fuel)                            %>% 
              dplyr::select(-c(time,ens_label,case_ens))           %>% 
              summarize_all(mean,na.rm=TRUE)                       %>% 
              ungroup()

sumfuel          = byfuel                                          %>%
                   group_by(case,ens_label,time)                   %>% 
                   dplyr::select(-fuel)                            %>% 
                   summarize_all(sum,na.rm=TRUE)                   %>% 
                   ungroup()                                       %>% 
                   mutate(month = month(time))                     %>% 
                   group_by(case,ens_label,month)                  %>% 
                   dplyr::select(-time)                            %>% 
                   summarize_all(mean,na.rm=TRUE)                  %>% 
                   ungroup()                                       %>% 
                   rename("fueload_byclass" = "fates_fuel_amount")

sumfuel          = sumfuel %>% left_join(fire_mmean,by=c("case","ens_label","month"))

fuel_plot = ggplot(sumfuel, aes(x=fueload_byclass,y=fates_fuel_amount,group=ens_label)) + geom_line(lwd=0.01) + facet_wrap(.~case,labeller=labeller(case=case_desc)) 
ggsave(fuel_plot,path=plot_main,filename="fueload-examine.jpeg",width=gg_width,height=gg_height,dpi=gg_depth)  
             
  
fire_loop = which(hlm1dvar$vnam %in% fire_vars)
gg_fire   = list()
for(h in fire_loop){
   h_vnam   = hlm1dvar$vnam[h] 
   h_desc   = hlm1dvar$desc [h]
   h_short  = hlm1dvar$short[h]
   h_unit   = hlm1dvar$unit [h]
   h_legend = v == 1
   cat0("   - ",h_desc,".")

  fire_now        = fire_mmean
   #mmean_now$case = factor(mmean_now$case,levels=unique(mmean_now$case))
   
   gg_now = ggplot(data=fire_now)
   gg_now = gg_now + facet_wrap(.~case,labeller=labeller(case=case_desc))
   gg_now = gg_now + geom_line(aes_string(x="month",y=h_vnam,group="ens_label",color="'Model'")
                              ,lwd=0.1,alpha=0.8,show.legend=TRUE)
   
   gg_now = gg_now + scale_colour_manual(name=""
                                        ,values=c("Model"= ens_colour[1]))
   gg_now = gg_now + labs(title=h_desc)
   gg_now = gg_now + scale_x_continuous( breaks = sequence(12))#end scale_x_continuous
   gg_now = gg_now + xlab(element_blank())
   gg_now = gg_now + ylab(desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz)
                         , axis.text.x        = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(h_vnam,"_monthlymean.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot      = gg_now
                    , device    = gg_device[d]
                    , path      = secy_path
                    , width     = gg_width
                    , height    = gg_height
                    , units     = gg_units
                    , dpi       = gg_depth
                    )
   }
  gg_fire[[h_vnam]] = gg_now      

}


# If sought, plot images on screen
if (gg_screen) gg_fire


```





```{r,label='load-site-data'}
if ("nc_site" %in% ls()){dummy = nc_close(nc_site); rm(nc_site)}

# TO-DO: loop for reading multiple site data
# Open NetCDF connection and retrieve variable names
cat0(" + Load site data from ",site_base,".")
nc_site  = nc_open(filename=site_file)
nc_nvars = nc_site$nvars
nc_ndims = nc_site$ndims
nc_dlist = rep(NA_character_,times=nc_ndims)
nc_vlist = rep(NA_character_,times=nc_nvars)
for (d in sequence(nc_ndims)) nc_dlist[d] = nc_site$dim[[d]]$name
for (v in sequence(nc_nvars)) nc_vlist[v] = nc_site$var[[v]]$name

# Select variables to load
nc_obs1d    = nc_vlist[tolower(nc_vlist) %in% hlm1dvar$vnam[hlm1dvar$assess]]
nc_obs2d    = nc_vlist[tolower(nc_vlist) %in% hlm2dsoi$vnam[hlm2dsoi$assess]]
nc_obs      = c(nc_obs1d, nc_obs2d)


# Extract time information
site_time0  = as_datetime(gsub(pattern="^days since ",replacement="",x=nc_site$dim$time$units))
site_time   = site_time0 + days(nc_site$dim$time$vals)
n_site_time = nc_site$dim$time$len

# Initialise a tibble that will host all data
site1d = tibble( time = site_time)

# Find conversion factors for monthly variables.
cmon.day = days_in_month(site1d$time)
cmon.hr  = day.hr  * cmon.day
cmon.min = day.min * cmon.day
cmon.sec = day.sec * cmon.day

# Loop through variables, and load data sets.
for (o in seq_along(nc_obs)){
   nc_nvnow        = nc_obs[o]
   if(nc_nvnow=="H2OSOI"){ #in hlm2dsoi
   h               = match(tolower(nc_nvnow),hlm2dsoi$vnam)
   h_vnam          = hlm2dsoi$vnam[h]
   h_desc          = hlm2dsoi$desc[h]
   h_add0          = eval(parse(text=hlm2dsoi$add0_sl[h]))
   h_mult          = eval(parse(text=hlm2dsoi$mult_sl[h]))
   nc_dat          = ncvar_get(nc=nc_site,varid=nc_nvnow)
   cat0("   - Retrieve ",h_desc,".")
   site1d[[h_vnam]] = h_add0 + h_mult * nc_dat  
   }else{
   h               = match(tolower(nc_nvnow),hlm1dvar$vnam)
   h_vnam          = hlm1dvar$vnam[h]
   h_desc          = hlm1dvar$desc[h]
   h_add0          = eval(parse(text=hlm1dvar$add0[h]))
   h_mult          = eval(parse(text=hlm1dvar$mult[h]))
   nc_dat          = ncvar_get(nc=nc_site,varid=nc_nvnow)
   cat0("   - Retrieve ",h_desc,".")
   site1d[[h_vnam]] = h_add0 + h_mult * nc_dat
   }
}#end for (o in seq_along(nc_obs))


# Load site coordinates. We will check whether observations and model are reasonably close.
site_coord = tibble( clon = c(ncvar_get(nc=nc_site,varid="LONGXY"))
                   , clat = c(ncvar_get(nc=nc_site,varid="LATIXY"))
                   , wlon = c(ncvar_get(nc=nc_site,varid="EDGEW" ))
                   , elon = c(ncvar_get(nc=nc_site,varid="EDGEE" ))
                   , slat = c(ncvar_get(nc=nc_site,varid="EDGES" ))
                   , nlat = c(ncvar_get(nc=nc_site,varid="EDGEN" ))
                   )#end tibble


# Close file and remove connection.
dummy = nc_close(nc_site)
rm(nc_site)

```






```{r,label='time-intersect'}
# Find the minimum interval that has overlaps with both site and model.
tstamp_each = unique(ens_tstamp)

tstampa_com = max(c(min(tstamp_each),min(site1d$time)))
tstampz_com = min(c(max(tstamp_each),max(site1d$time)))

# Restrict time to the overlapping period
tstamp_com  = tstamp_each[(tstamp_each >= tstampa_com) & (tstamp_each <= tstampz_com)]
ntstamp_com = length(tstamp_com)

# Trim site observations to the overlapping period. Also, standardize the data so we
# use only one of NA or NaN.
site = site1d %>%
         filter( (time >= tstampa_com) & (time <= tstampz_com)) %>%
         mutate_at(vars(-time),function(x) ifelse(is.nan(x),NA,x))

#  Load site LAI and AGB to bind with site
lai = fread(site_lai)
lai = lai                                %>% 
      mutate(LAI_DATE = ymd(LAI_DATE)
            ,year     = year(LAI_DATE)
            ,month    = month(LAI_DATE)) %>% 
      dplyr::select(-LAI_DATE)           %>% 
      group_by(year,month)               %>% 
      summarize_all(mean,na.rm=TRUE)     %>% 
      filter(!is.na(year))               

colnames(lai) <- c("year","month","elai","elai_sd")
agb = fread(site_agb)
agb = agb                                 %>% 
      mutate(date       = ymd(date)
            ,year       = year(date)
            ,month      = month(date)
            ,biomass    = biomass/2
            ,biomass_sd = biomass_sd/2)   %>% 
      dplyr::select(-date)                %>% 
      group_by(year,month)                %>% 
      summarise_all(mean,na.rm=TRUE)      %>% 
      ungroup()
colnames(agb) <- c("year","month","agb","agb_sd","sample")

## combine site observations
site  = site                                  %>%  
        mutate(year  = year(time)
              ,month = month(time))           %>% 
        left_join(agb,by=c("year","month"))   %>% 
        left_join(lai,by=c("year","month"))   %>% 
        dplyr::select(-c(agb_sd,sample,elai_sd))

site_var = names(site)
site_var = gsub(pattern="gpp",replacement="fates_gpp",x=site_var)
site_var = gsub(pattern="agb",replacement="fates_vegc_aboveground",x=site_var)
names(site) = site_var


```







```{r,label='merge-data',message=FALSE,results='hide'}
# First we have to filter out HLM variables that are not available at site or within the time frame of obs; 
# and then combine HLM with site obs to make comparison plots

cat0(" + Find the nearest grid cell from model simulation.")

site_xy = site_coord [1:2] %>% mutate(clon=clon-360)
nn_pt   = RANN::nn2(orig_xy[,1:2],site_xy[,1:2],k=1)
xy_model= orig_xy[nn_pt$nn.idx]
xy_model = xy_model %>% mutate(lon=round(lon,5),lat=round(lat,5))

    
hlm_site = hlm1d                                              %>% 
           mutate(lon=round(lon,5)
                 ,lat=round(lat,5))                           %>% 
           filter(lon == xy_model$lon & lat == xy_model$lat)   
   
          

# intersect variables
cat0(" + Find the variables common to both site and model.")
# Find common variables
var_both = intersect(names(site),names(hlm_site))
var_both = var_both[! var_both %in% "time"]
var_both = var_both[c(3,4,6,8,9)] # only select 5

# Restrict variables for both site and model
site  = site %>% dplyr::select( all_of(c("time",var_both)))

hlm_site = hlm_site                                                      %>% 
           dplyr::select( all_of(c('time','case',var_both)))  

cat0(" + Merge data sets into a single tibble.")

emean_com  = rbind( site %>% mutate(case = 0L)
                  , hlm_site)   %>%
             rename(source = case)


# Find the mean seasonal cycle.
mmean_com = emean_com                                                     %>%
            mutate(month = month(time))                                   %>%
            group_by(month,source)                                        %>%
            dplyr::select(-time)                                          %>%
            summarise_all(mean, na.rm=TRUE)                               %>%
            ungroup()                                                     %>%
            rename_at(vars(var_both), function(x) paste0(x,"_mean"))

# Find the lower range of the seasonal cycle
mqlwr_com = emean_com                                                         %>%
            mutate(month = month(time))                                       %>%
            group_by(month,source)                                            %>%
            dplyr::select(-time)                                              %>%
            summarise_all(quantile,probs=qlwr_ribbon,names=FALSE, na.rm=TRUE) %>%
            ungroup()                                                         %>%
            rename_at(vars(var_both), function(x) paste0(x,"_qlwr"))

# Find the upper range of the seasonal cycle
mqupr_com = emean_com                                                         %>%
            mutate(month = month(time))                                       %>%
            group_by(month,source)                                            %>%
            dplyr::select(-time)                                              %>%
            summarise_all(quantile,probs=qupr_ribbon,names=FALSE, na.rm=TRUE) %>%
            ungroup()                                                         %>%
            rename_at(vars(var_both), function(x) paste0(x,"_qupr"))

# Find the annual averages.
mmean_com = as_tibble( merge( x  = merge(x=mmean_com,y=mqlwr_com,by=c("month","source"))
                        , y  = mqupr_com
                        , by = c("month","source")
                        )#end merge
                 )#end as_tibble

# Sort data by month and source
mmean_com = mmean_com %>% arrange(source,month)         


```








```{r,label='model-site-comparison',message=FALSE,results='hide',fig.width=15, fig.height=12}
cat0(" + Plot the mean seasonal cycle of site measurements/estimates and model predictions.")

# Set legends
leg_colours = c(site_colour,ens_colour)
#leg_labels  = c("Site",as.character(case_desc))

gg_emean = list()
for( n in sequence(ncase)){
  case_now = case_names[n]
  desc_now = case_desc[[n]]
  h_mmean        = mmean_com %>% filter(source%in%c(0,case_now))
  leg_labels      = c("Site",desc_now)
  for (v in seq_along(var_both)){
   # Load variable
   h        = c(match(var_both[v],hlm1dvar$vnam), match(var_both[v],hlm2dsoi$vnam))
   hvars = c("hlm1dvar", "hlm2dsoi")
   whichvar = hvars[which(!is.na(h))]
   var_now  = get(whichvar)
   h        =h[!is.na(h)]
   h_vnam   = var_now$vnam [h]
   h_vmean  = paste0(h_vnam,"_mean")
   h_vqlwr  = paste0(h_vnam,"_qlwr")
   h_vqupr  = paste0(h_vnam,"_qupr")
   h_desc   = var_now$desc [h]
   h_short  = var_now$short[h]
   h_unit   = var_now$unit [h]
   h_legend = v == 1 


   # Temporary data table. We convert the classes back to factor.

   h_mmean$source = factor(h_mmean$source,levels=unique(h_mmean$source))

   # Initialise plot (we use line)
   gg_now = ggplot( data    = h_mmean
                  , mapping = aes_string( x      = "month"
                                        , group  = "source"
                                        , colour = "source"
                                        , fill   = "source"
                                        )
                  )#end ggplot
   gg_now = gg_now + scale_colour_manual(name="",aesthetics="colour",labels=leg_labels,values=leg_colours)
   gg_now = gg_now + scale_colour_manual(name="",aesthetics="fill"  ,labels=leg_labels,values=leg_colours)

   # We only add legend for a single plot.  Patchwork will fix this in the end.
   gg_now = gg_now + geom_ribbon( aes_string(ymin=h_vqlwr,ymax=h_vqupr)
                                , alpha       = alpha_ribbon
                                , show.legend = h_legend
                                , colour      = "transparent"
                                )#end geom_ribbon
   gg_now = gg_now + geom_line( aes_string(y=h_vmean)
                              , linewidth = 0.7
                              , show.legend = h_legend
                              )#end geom_line

   
   # Add local annotation
   gg_now = gg_now + labs(title=element_blank())
   gg_now = gg_now + scale_x_continuous( breaks = sequence(12)
                                       , labels = substring(month.abb,1,1)
                                       )#end scale_x_continuous
   gg_now = gg_now + xlab(element_blank())
   gg_now = gg_now + ylab(desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))
   gg_now = gg_now + theme_bw( base_size      = gg_ptsz
                             , base_family    = "Helvetica"
                             , base_line_size = 0.5
                             , base_rect_size = 0.5
                               )#end theme_grey

   # Additional legend settings
   # if (h_legend) gg_now = gg_now + theme( legend.position = "bottom")

   # Axis settings
   gg_now = gg_now + theme( axis.text.x       = element_text( size   = gg_ptsz
                                                            , margin = unit(rep(0.35,times=4),"char")
                                                            )#end element_text
                          , axis.text.y       = element_text( size   = gg_ptsz
                                                            , margin = unit(rep(0.35,times=4),"char")
                                                            )#end element_text
                          , axis.ticks.length = unit(-0.2,"char")
                          , axis.title.y      = element_text( size = gg_ptsz * 0.9)
                          )#end theme

  # Write plot settings to the list.
  gg_emean[[h_vnam]] = gg_now
}#end for (v in seq_along(var_both)){

# Wrap plots then global settings and x axis.
gg_patch = wrap_plots(gg_emean)
gg_patch = gg_patch + guide_area() + plot_layout(guides="collect")
gg_patch = gg_patch + plot_annotation(tag_levels = "a")

# Save plots.
for (d in sequence(ndevice)){
   h_output = paste0("comparison_meanseason",case_now,".",gg_device[d])
   dummy    = ggsave( filename = h_output
                    , plot     = gg_patch
                    , device   = gg_device[d]
                    , path     = secy_path
                    , width    = gg_width
                    , height   = gg_height
                    , units    = gg_units
                    , dpi      = gg_depth
                    )#end ggsave

}#end for (d in sequence(n_device))
  
}


# If sought, plot images on screen
if (gg_screen) gg_patch


```






```{r,label="double-check-agb-seabloom-data"}

## AGB from Vaira Ranch likely includes both live and dead mass, we double check with
## Seabloom et al. 2020 data https://portal.edirepository.org/nis/mapbrowse?packageid=edi.599.2
## at Sierra foothill site
seablm_fates = survi_amean                                   %>% 
               dplyr::select(case,ens_label,case_ens
                            ,year,fates_vegc_aboveground)    %>% 
               rename(agb = fates_vegc_aboveground) 
keep_fates   = survi_amean                                   %>% 
               dplyr::select(case,ens_label,case_ens
                            ,year,fates_vegc_aboveground)    %>% 
               filter(case_ens %in% ens_keep)                %>% 
               rename(agb = fates_vegc_aboveground) 
#amean_now = seablm_fates
amean_now  = keep_fates
amean_now$case = factor(amean_now$case,levels=unique(amean_now$case))
   
gg_now = ggplot(data=amean_now)
gg_now = gg_now + geom_line(aes(x=year,y=agb,group=ens_label,color="Model")
                              ,lwd=1,alpha=0.8,show.legend=TRUE)
gg_now = gg_now + geom_line(data = sierra_amean,aes(x=year,y=agb,color="Seabloom-SierraFoothill"))
gg_now = gg_now + scale_colour_manual(name="",values=c("Model"=ens_colour[1]
                                                ,"Seabloom-SierraFoothill"=site_colour))
gg_now = gg_now + geom_ribbon( data = sierra_amean
                                , aes(x = year,ymin=agb_qlwr,ymax=agb_qupr)
                                , alpha       = alpha_ribbon
                                , inherit.aes = FALSE
                                , show.legend = h_legend
                                , colour      = "transparent"
                                )
gg_now = gg_now + facet_wrap(.~case,labeller=labeller(case=case_desc))    

gg_now = gg_now + xlab("Year")
gg_now = gg_now + ylab(expression(Aboveground~biomass~"("~kgC~m^{-2}~")"))
gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
gg_now = gg_now + theme(axis.text.x        = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         , axis.title.y = element_text(size = gg_ptsz*0.9)
                         , legend.text=element_text(size=gg_ptsz*0.7)
                         , plot.title = element_text(size = gg_ptsz*0.95)
                         ) #end theme
p_output = paste0("seablm-keep-agb.",gg_device)
dummy    = ggsave( filename = p_output
                 , plot     = gg_now
                 , device   = gg_device[d]
                 , path     = plot_main
                 , width    = gg_width
                 , height   = gg_height*0.7
                 , units    = gg_units
                 , dpi      = gg_depth
                    )

if(gg_screen) gg_now

```






```{r,label="test-unit}

test = hlm1d   %>%  dplyr::select(case,lon,lat,time,fates_burnfrac)
test = test %>% filter(case==1)
test = test %>% mutate(lon = round(lon,digits = 5),
                       lat = round(lat,digits=5))
test11 = test 
test11$lon_f = factor(test11$lon,levels=unique(test11$lon))
test11$lat_f = factor(test11$lat,levels=unique(test11$lat))

test11 = test11 %>% group_by(case,lon_f,lat_f)   %>% dplyr::select(-time)  %>% mutate(bfrac=mean(fates_burnfrac)) %>% ungroup() %>% 
  select(lon,lat,bfrac) %>% distinct()
test2  = test %>% filter(time=="2020-09-01")

gpp      = ncvar_get(nc_conn,"FATES_GPP")
bfrac    = ncvar_get(nc_conn,"FATES_BURNFRAC")

gpp_vec  = gpp[,,472] #2020-04
gpp_vec  = matrix(gpp_vec,ncol=1)
gpp_vec  =  gpp_vec*2592000

bf_vec   = bfrac[,,476] #2020-08
bf_vec   = matrix(bf_vec,ncol=1)

fil_var  = matrix(test11$bfrac, nrow=147,ncol=151)
x_arr    = matrix(test11$lon,nrow=147,ncol=151)
y_arr    = matrix(test11$lat,nrow=147,ncol=151)
wrf_star = st_as_stars(fil_var)
wrf_star = st_as_stars(wrf_star, curvilinear=list(X1=x_arr,X2=y_arr),crs=wgs)
wrf_sf   = st_as_sf(wrf_star,na.rm=FALSE)
ca_co    = USAboundaries::us_counties(resolution = "high", states = "CA")

##plot to see how the active domain looks like

p = ggplot() + geom_sf(data=wrf_sf,colour="grey50", aes(fill=A1),lwd=0)+
  coord_sf(crs=st_crs(wgs)) + 
  my_theme +scale_fill_continuous(low="thistle2", high="darkred", 
                                  guide="colorbar",na.value="grey50")+
  geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) +
  geom_point(aes(x=-120.9508,y=38.4133), colour=alpha("blue",0.6), size=0.2)+
  labs(x="",y="") +
  guides(fill=guide_legend(title="Burned fraction"))
p

ggsave(p, filename="case1-bf.jpeg",path=secy_path,width=gg_widthn,height=gg_height,dpi=gg_depth,device=gg_device[1])



fil_var  = matrix(mar_rain$rain, nrow=147,ncol=151)
x_arr    = matrix(mar_rain$lon,nrow=147,ncol=151)
y_arr    = matrix(mar_rain$lat,nrow=147,ncol=151)
wrf_star = st_as_stars(fil_var)
wrf_star = st_as_stars(wrf_star, curvilinear=list(X1=x_arr,X2=y_arr),crs=wgs)
wrf_sf   = st_as_sf(wrf_star,na.rm=FALSE)
ca_co    = USAboundaries::us_counties(resolution = "high", states = "CA")

##plot to see how the active domain looks like

p = ggplot() + geom_sf(data=wrf_sf,colour="grey50", aes(fill=A1),lwd=0)+
  coord_sf(crs=st_crs(wgs)) + 
  my_theme +scale_fill_continuous(low="thistle2", high="darkred", 
                                  guide="colorbar",na.value="grey50")+
  geom_sf(data = ca_co, color = alpha("black", alpha=1),lwd=0.1,fill=NA) +
  geom_point(aes(x=-120.9508,y=38.4133), colour=alpha("blue",0.6), size=0.2)+
  labs(x="",y="") +
  guides(fill=guide_legend(title="Precip-March"))
p

ggsave(p, filename="case1-bf.jpeg",path=secy_path,width=gg_widthn,height=gg_height,dpi=gg_depth,device=gg_device[1])

```

